---
title: "TPA - P:R analysis"
author: "Hugo Scharfenstein"
date: "`r Sys.Date()`"
output: html_document
---

Load packages
```{r}
#Data import and export packages
library(readxl)
library(writexl)
#Data wrangling packages
library(tidyverse)
#Respiration analysis package
library(respR)
```

Import data
```{r}
#Import raw O2 progression data into R
PR.df <- read_xlsx("./GCB dataset.xlsx", sheet = 7) %>%
  mutate(phase = Phase) %>%
  mutate(time = Time_seconds) %>%
  mutate(oxygen = O2) %>%
  filter(time > 10 ) %>% #The first 10 seconds are filtered out due to O2 values adjusting after starting measurements
  select(Growth_temperature, EE_treatment, Run, Chamber, Lineage, Replicate, phase, time, oxygen)

#Analyse one temperature at a time
PR.df <- PR.df %>%
  filter(Growth_temperature == 21) #Change temperature here to toggle between data sets

#Prepare labels for plot facets
growth.labs <- c("Growth cycle 1", "Growth cycle 2")
names(growth.labs) <- c('1', "2")
temperature.labs <- c('21°C', '23°C', '25°C', '27°C', '29°C', '31°C', '33°C')
names(temperature.labs) <- c('21', '23', '25', '27', '29', '31', '33')
treatment.labs <- c('Fluc-short', 'Fluc-med', 'Fluc-long', 'Cont-ele', 'Cont-amb')
names(treatment.labs) <- c('TP1', 'TP2', 'TP3', 'TP4', 'TP5')
```

Inspect raw O2 measurements
```{r}
#Filter for dark phase to visualize respiration rates
Rd.plot <- PR.df %>%
  filter(phase == 'dark') %>%
  ggplot(aes(x = time, y = oxygen, colour = EE_treatment)) +
  geom_hline(aes(yintercept = 80), colour="black", size = 0.5, linetype = 'dashed') +
  geom_vline(aes(xintercept = 900), colour="black", size = 0.5, linetype = 'solid') +
  geom_vline(aes(xintercept = 1200), colour="black", size = 0.5, linetype = 'solid') +
  geom_point(size = 0.1) +
  geom_point(data = filter(PR.df, Lineage == 'Blank' & phase == 'dark'), colour="black", size = 0.1) +
  facet_grid(EE_treatment~., labeller = labeller(EE_treatment = treatment.labs)) +
  scale_x_continuous('Time (seconds)', limits = c(0, 1210), breaks = seq(0, 1210, by = 400), expand = c(0,0)) +
  scale_y_continuous('Oxygen (% air saturation)') +
  scale_colour_manual(values=c("#8B0000","#E4B031", "#26727B", "#4770B3", "#58595b")) +
  theme_bw() +
  theme(panel.grid.minor = element_blank(), 
        strip.text.x=element_text(size=14, face="bold"),
        strip.text.y=element_text(size=14, face="bold"),
        title = element_text(size = 14, face = "bold"),
        legend.title = element_text(size = 14, face = "bold"),
        legend.text = element_text(size = 12),
        legend.position = 'top',
        axis.line = element_line(colour = "black"),
        axis.title.x = element_text(size = 14, colour = "black", face = "bold"),
        axis.text.x = element_text(size = 12, colour = "black"), 
        axis.title.y = element_text(size = 14, colour = "black", face = "bold"), 
        axis.text.y = element_text(size = 12, colour = "black")) +
  ggtitle(paste0("Change in oxygen at ", PR.df$Growth_temperature, "°C - dark phase")) +
  guides(colour = FALSE)
Rd.plot

#Filter for light phase to visualize photosynthesis rates
Pnet.plot <- PR.df %>%
  filter(phase == 'light') %>%
  ggplot(aes(x = time, y = oxygen, colour = EE_treatment)) +
  geom_hline(aes(yintercept = 80), colour="black", size = 0.5, linetype = 'dashed') +
  geom_vline(aes(xintercept = 1500), colour="black", size = 0.5, linetype = 'solid') +
  geom_vline(aes(xintercept = 2100), colour="black", size = 0.5, linetype = 'solid') +
  geom_point(size = 0.05) +
  geom_point(data = filter(PR.df, Lineage == 'Blank' & phase == 'light'), colour="black", size = 0.05) +
  facet_grid(EE_treatment~., labeller = labeller(EE_treatment = treatment.labs)) +
  scale_x_continuous('Time (seconds)', limits = c(1200, 2200), breaks = seq(1200, 2200, by = 400), expand = c(0,0)) +
  scale_y_continuous('Oxygen (% air saturation)') +
  scale_colour_manual(values=c("#8B0000","#E4B031", "#26727B", "#4770B3", "#58595b")) +
  theme_bw() +
  theme(panel.grid.minor = element_blank(), 
        strip.text.x=element_text(size=14, face="bold"),
        strip.text.y=element_text(size=14, face="bold"),
        title = element_text(size = 14, face = "bold"),
        legend.title = element_text(size = 14, face = "bold"),
        legend.text = element_text(size = 12),
        legend.position = 'top',
        axis.line = element_line(colour = "black"),
        axis.title.x = element_text(size = 14, colour = "black", face = "bold"),
        axis.text.x = element_text(size = 12, colour = "black"), 
        axis.title.y = element_text(size = 14, colour = "black", face = "bold"), 
        axis.text.y = element_text(size = 12, colour = "black")) +
  ggtitle(paste0("Change in oxygen at ", PR.df$Growth_temperature, "°C - light phase")) +
  guides(colour = FALSE)
Pnet.plot
```
Mark samples that displayed irregularities during the light phase (chambers 6,7 and 9 appear to have technical issues)
```{r}
PR.df_QC <- PR.df %>%
  #mutate(QC = case_when(Lineage %in% c('TP1-3','TP2-2', 'TP3-2', 'TP4-3', 'TP5-1') ~ 'throw',
                        #T ~ 'keep')) #For 31°C
  #mutate(QC = case_when(Lineage %in% c('TP1-2', 'TP2-1', 'TP4-3', 'TP4-5', 'TP5-1', 'TP5-6') ~ 'throw',
                        #T ~ 'keep')) #For 29°C
  #mutate(QC = case_when(Lineage %in% c('TP2-6') ~ 'throw',
                        #T ~ 'keep')) #For 27°C
  #mutate(QC = case_when(Lineage %in% c('TP1-4', 'TP2-5', 'TP3-6', 'TP4-4', 'TP5-6') ~ 'throw',
                        #T ~ 'keep')) #For 25°C
  #mutate(QC = case_when(Lineage %in% c('TP2-2', 'TP2-3', 'TP5-4', 'TP5-6') ~ 'throw',
                        #T ~ 'keep')) #For 23°C
  mutate(QC = case_when(Lineage %in% c('TP2-2', 'TP2-6', 'TP3-2') ~ 'throw',
                        T ~ 'keep')) #For 21°C
```

Plot change in O2 over the entire assay, indicating samples that failed QC
```{r}
PR.plot <- PR.df_QC %>%
  ggplot(aes(x = time, y = oxygen, colour = EE_treatment)) +
  annotate('rect', xmin = 0, xmax = 1200, ymin = 75, ymax = 120, alpha = 0.2, fill = 'black') +
  geom_hline(aes(yintercept = 80), colour="black", size = 0.25, linetype = 'dashed') +
  geom_vline(aes(xintercept = 900), colour="black", size = 0.5, linetype = 'solid') +
  geom_vline(aes(xintercept = 1200), colour="black", size = 0.5, linetype = 'solid') +
  geom_vline(aes(xintercept = 1500), colour="black", size = 0.5, linetype = 'solid') +
  geom_vline(aes(xintercept = 2100), colour="black", size = 0.5, linetype = 'solid') +
  geom_point(aes(alpha = QC), size = 0.001) +
  geom_point(data = filter(PR.df_QC, Lineage == 'Blank' & Replicate == 1), colour="black", size = 0.01) +
  facet_grid(EE_treatment~., labeller = labeller(EE_treatment = treatment.labs)) +
  scale_alpha_manual(values = c(0.75, 0.05)) +
  scale_x_continuous('Time (seconds)', limits = c(0, 2200), breaks = seq(0, 2200, by = 400), expand = c(0,0)) +
  scale_y_continuous('Oxygen (% air saturation)', limits = c(75, 120), expand = c(0,0)) +
  scale_colour_manual(values=c("#8B0000","#E4B031", "#26727B", "#4770B3", "#58595b")) +
  theme_bw() +
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(), 
        strip.text.x=element_text(size=14, face="bold"),
        strip.text.y=element_text(size=14, face="bold"),
        title = element_text(size = 14, face = "bold"),
        legend.title = element_text(size = 14, face = "bold"),
        legend.text = element_text(size = 12),
        legend.position = 'top',
        axis.line = element_line(colour = "black"),
        axis.title.x = element_text(size = 14, colour = "black", face = "bold"),
        axis.text.x = element_text(size = 12, colour = "black"), 
        axis.title.y = element_text(size = 14, colour = "black", face = "bold"), 
        axis.text.y = element_text(size = 12, colour = "black")) +
  ggtitle(paste0("Change in oxygen at ", PR.df_QC$Growth_temperature, "°C")) +
  guides(colour = F, alpha = F)
PR.plot
```

Save and export plots
```{r}
#PR.plot_31C <- PR.plot
#PR.plot_29C <- PR.plot
#PR.plot_27C <- PR.plot
#PR.plot_25C <- PR.plot
#PR.plot_23C <- PR.plot
PR.plot_21C <- PR.plot

PR_plot <- ggarrange(PR.plot_31C + rremove('xlab') + rremove('ylab'),
                     PR.plot_29C + rremove('xlab') + rremove('ylab'),
                     PR.plot_27C +rremove('xlab') + rremove('ylab'),
                     #PR.plot_25C +rremove('xlab') + rremove('ylab'),
                     #PR.plot_23C +rremove('xlab') + rremove('ylab'),
                     #PR.plot_21C +rremove('xlab') + rremove('ylab'),
                     common.legend = TRUE,
                     heights = c(1, 1, 1),
                     widths = c(1, 1, 1),
                     ncol = 3) 
PR_plot

PR_plot.final <- annotate_figure(PR_plot,
                                 left = text_grob("Oxygen (% air saturation)", color = "black", face = "bold", rot = 90, size = 14),
                                 bottom = text_grob("Time (seconds)", color = "black", face = "bold", size = 14))
PR_plot.final

ggsave(plot = PR_plot.final, "TPA - raw O2 measurements.pdf", 
       path = "./",
       width = 12.5,
       height = 7.5,
       units = 'in',
       dpi = 300)
```

Respiration analysis

Inspect data
```{r}
#Filter for dark phase to visualize respiration
Rd.plot <- PR.df %>%
  filter(phase == 'dark') %>%
  ggplot(aes(x = time, y = oxygen, colour = EE_treatment)) +
  geom_hline(aes(yintercept = 80), colour="black", size = 0.5, linetype = 'dashed') +
  geom_vline(aes(xintercept = 900), colour="black", size = 0.5, linetype = 'solid') +
  geom_vline(aes(xintercept = 1200), colour="black", size = 0.5, linetype = 'solid') +
  geom_point(size = 0.1) +
  geom_point(data = filter(PR.df, Lineage == 'Blank' & phase == 'dark'), colour="black", size = 0.1) +
  facet_grid(.~EE_treatment, labeller = labeller(EE_treatment = treatment.labs)) +
  scale_x_continuous('Time (seconds)', limits = c(0, 1210), breaks = seq(0, 1210, by = 400), expand = c(0,0)) +
  scale_y_continuous('Oxygen (% air saturation)') +
  scale_colour_manual(values=c("#8B0000","#E4B031", "#26727B", "#4770B3", "#58595b")) +
  theme_bw() +
  theme(panel.grid.minor = element_blank(), 
        strip.text.x=element_text(size=14, face="bold"),
        strip.text.y=element_text(size=14, face="bold"),
        title = element_text(size = 14, face = "bold"),
        legend.title = element_text(size = 14, face = "bold"),
        legend.text = element_text(size = 12),
        legend.position = 'top',
        axis.line = element_line(colour = "black"),
        axis.title.x = element_text(size = 14, colour = "black", face = "bold"),
        axis.text.x = element_text(size = 12, colour = "black"), 
        axis.title.y = element_text(size = 14, colour = "black", face = "bold"), 
        axis.text.y = element_text(size = 12, colour = "black")) +
  ggtitle(paste0("Change in oxygen at ", PR.df$Growth_temperature, "°C - dark phase")) +
  guides(colour = F)
Rd.plot
```

Calculate respiration rates
```{r}
#Separate culture from blank measurements
Rd.df <- PR.df %>%
  filter(phase == 'dark') %>%
  filter(Lineage != 'Blank') %>%
  select(!c(Replicate, Run, Chamber, phase)) %>%
  nest(data = c(time, oxygen))

#Prepare the function calc_nest_rate which enables us to map the function calc_rate() to each nest
calc_nest_rate <- function(nest_data) {
  result <- calc_rate(data.frame(time = nest_data$time, oxygen = nest_data$oxygen),
            from = 900, to = 1200, by = 'time' #Calculate rates for last 5 minutes of the dark phase (first 15 mins were for dark adaptation)
            )
  list(intercept = result$summary[[3]], slope = result$summary[[4]], r2 = result$summary[[5]], rate = result$summary[[13]]) #Hash this out to obtain full list output
}

#Prepare the function extract_values to extract values of interest from the generated lists
extract_values_resp <- function(nested_list) {
  data.frame(intercept = nested_list$intercept,
             slope = nested_list$slope,
             r2 = nested_list$r2,
             rate = nested_list$rate)
}

#Calculate respiration rates
Rd.df_rate <- Rd.df %>%
  mutate(rate_lists = map(data, calc_nest_rate)) %>%
  mutate(rate_values = map(rate_lists, extract_values_resp)) %>%
  select(-rate_lists) %>%
  unnest(rate_values) %>%
  relocate(data, .after = rate) #%>%
  #filter(Lineage != 'TP1-1') #For time series at 31C, TP1-1 resp rate needs to be calculated between 5 & 10 mins

#For time series at 31C, calculate respiration rate for TP1-1 between 5-10 mins
Rd.df_rate_TP1.1_31C <- Rd.df %>%
  filter(Lineage == 'TP1-1') %>%
  mutate(rate_lists = map(data, calc_nest_rate)) %>%
  mutate(rate_values = map(rate_lists, extract_values_resp)) %>%
  select(-rate_lists) %>%
  unnest(rate_values) %>%
  relocate(data, .after = rate)

Rd.df_rate <- rbind(Rd.df_rate, Rd.df_rate_TP1.1_31C)
```

Calculate & adjust for background respiration rates using blanks
```{r}
#Select blank measurements
Rd.df_blank <- PR.df %>%
  filter(phase == 'dark') %>%
  filter(Lineage == 'Blank') %>%
  filter(EE_treatment != 'TP1' | Replicate != 2) %>% #TP1 usually has a 2 technical replicates for the blank, use either
  select(!c(Replicate, Run, Chamber, phase)) %>%
  nest(data = c(time, oxygen))

#Calculate background respiration rates
Rd.df_blank_rate <- Rd.df_blank %>%
  mutate(rate_lists = map(data, calc_nest_rate)) %>%
  mutate(rate_values = map(rate_lists, extract_values_resp)) %>%
  select(-rate_lists) %>%
  unnest(rate_values) %>%
  relocate(data, .after = rate)

#Adjust for background respiration rates
Rd.df_adj <- rbind(Rd.df_rate, Rd.df_blank_rate) %>%
  group_by(EE_treatment) %>%
  mutate(adj_rate = rate - rate[Lineage == 'Blank']) %>%
  relocate(data, .after = adj_rate)
```

Convert the adjusted respiration rates to pmolO2/min/cell
```{r}
#Convert to pmolO2/min
Rd_conversion <- convert_rate(Rd.df_adj$adj_rate, 
                        oxy.unit = '%Air',           # oxygen unit to convert from
                        time.unit = 's',             # time unit to convert from
                        output.unit = 'pmol/min',    # unit to convert to
                        volume = 0.00146,            # chamber volume in litres (1.46 ml)
                        t = 21,                      # temperature, in °C (change for each temperature treatment)
                        S = 33,                      # salinity, in ppt
                        P = 1.013253)                # atmospheric pressure, in bar                   

Rd.df_converted <- list(adj_rate = Rd_conversion$summary[[13]], pmolO2.s = Rd_conversion$summary[[29]]) %>%
  as.data.frame()

Rd.df_final <- left_join(Rd.df_adj, Rd.df_converted, by = 'adj_rate') %>%
  distinct() %>%
  mutate(pmolO2.min = pmolO2.s*60) %>% #convert from pmol/s to pmol/min
  mutate(pmolO2.min.cell = pmolO2.min/(750000*1.46)) #normalise for cell density (750,000 cells/ml)
```

Visualize Rd rates
```{r}
Rd.plot_final <- Rd.df_final %>%
  filter(Lineage != 'Blank') %>%
  ggplot(aes(y = pmolO2.min.cell, x = EE_treatment, fill = EE_treatment)) +
  geom_hline(aes(yintercept = 0), colour="black", size = 0.5, linetype = 'dashed') +
  geom_boxplot() +
  geom_point() +
  scale_x_discrete(labels = treatment.labs, 'Experimental evolution treatment') +
  scale_y_continuous(expression(bold(R[d]~(pmol~O[2]~min^-1~cell^-1)))) +
  scale_fill_manual(values=c("#8B0000","#E4B031", "#26727B", "#4770B3", "#58595b")) +
  theme_bw() +
  theme(panel.grid.major.x = element_blank(),
        panel.grid.minor = element_blank(),
        strip.text.x=element_text(size=14, face="bold"),
        strip.text.y=element_text(size=14, face="bold"),
        plot.title = element_text(size = 14, face = "bold"),
        legend.text = element_text(size = 12),
        axis.line = element_line(colour = "black"),
        axis.title.x = element_text(size = 14, colour = "black", face = "bold"),
        axis.text.x = element_text(size = 12, angle = 45, hjust = 1, colour = "black"), 
        axis.title.y = element_text(size = 14, colour = "black", face = "bold"), 
        axis.text.y = element_text(size = 12, colour = "black")) +
  ggtitle(paste0("Rd rates at ", Rd.df_final$Growth_temperature, "°C")) +
  guides(fill = FALSE, colour = FALSE)
Rd.plot_final
```

Save respiration rates
```{r}
Rdf_xl <- Rdf_final %>%
  select(-data) %>%
  arrange(by_group = Lineage) %>%
  write_xlsx(path = "./Rd rates - 21C.xlsx", #For each temperature, change file name
             col_names = TRUE,
             format_headers = TRUE)
```

Photosynthesis rates

Inspect data
```{r}
#Filter for light phase to visualize photosynthesis
Pnet.plot <- PR.df %>%
  filter(phase == 'light') %>%
  #filter(Lineage == 'TP4-3') %>%
  ggplot(aes(x = time, y = oxygen, colour = EE_treatment)) +
  geom_hline(aes(yintercept = 80), colour="black", size = 0.5, linetype = 'dashed') +
  geom_vline(aes(xintercept = 1500), colour="black", size = 0.5, linetype = 'solid') +
  geom_vline(aes(xintercept = 2100), colour="black", size = 0.5, linetype = 'solid') +
  geom_point(size = 0.05) +
  geom_point(data = filter(PR.df, Lineage == 'Blank' & phase == 'light'), colour="black", size = 0.05) +
  facet_grid(.~EE_treatment, labeller = labeller(EE_treatment = treatment.labs)) +
  scale_x_continuous('Time (seconds)', limits = c(1200, 2200), breaks = seq(1200, 2200, by = 400), expand = c(0,0)) +
  scale_y_continuous('Oxygen (% air saturation)') +
  scale_colour_manual(values=c("#8B0000","#E4B031", "#26727B", "#4770B3", "#58595b")) +
  theme_bw() +
  theme(panel.grid.minor = element_blank(), 
        strip.text.x=element_text(size=14, face="bold"),
        strip.text.y=element_text(size=14, face="bold"),
        title = element_text(size = 14, face = "bold"),
        legend.title = element_text(size = 14, face = "bold"),
        legend.text = element_text(size = 12),
        legend.position = 'top',
        axis.line = element_line(colour = "black"),
        axis.title.x = element_text(size = 14, colour = "black", face = "bold"),
        axis.text.x = element_text(size = 12, colour = "black"), 
        axis.title.y = element_text(size = 14, colour = "black", face = "bold"), 
        axis.text.y = element_text(size = 12, colour = "black")) +
  ggtitle(paste0("Change in oxygen at ", PR.df$Growth_temperature, "°C - light phase")) +
  guides(colour = F)
Pnet.plot
```

Calculate Pnet rates
```{r}
#Separate culture from blank measurements
Pnet.df <- PR.df %>%
  filter(phase == 'light') %>%
  filter(Lineage != 'Blank') %>%
  #filter(!(Lineage %in% c('TP1-3', 'TP2-2', 'TP3-2', 'TP4-3', 'TP5-1'))) %>% #Remove outliers at 31°C
  #filter(!(Lineage %in% c('TP1-2', 'TP2-1', 'TP4-3', 'TP4-5', 'TP5-1', 'TP5-5'))) %>% #Remove outliers at 29°C
  #filter(!(Lineage %in% c('TP2-6'))) %>% #Remove outliers at 27°C
  #filter(!(Lineage %in% c('TP1-4', 'TP2-5', 'TP3-6', 'TP4-4', 'TP5-6'))) %>% #Remove outliers at 25°C
  #filter(!(Lineage %in% c('TP2-2', 'TP2-3', 'TP5-4', 'TP5-6'))) %>% #Remove outliers at 23°C
  #filter(!(Lineage %in% c('TP2-2', 'TP2-6', 'TP3-2'))) %>% #Remove outliers at 21°C
  filter(time > 1500) %>% #The first 5 mins are filtered due to O2 values adjusting after turning on lights
  filter(time < 2100) %>% #Only keep the following 10 mins of light phase to calculate photosynthesis rates
  select(!c(Replicate, Run, Chamber, phase)) %>%
  nest(data = c(time, oxygen))

#Prepare the function auto_nest_rate which enables us to map the function auto_rate() to each nest
auto_nest_rate <- function(nest_data) {
  result <- auto_rate(data.frame(time = nest_data$time, oxygen = nest_data$oxygen),
            width = 300, by = "time"
            )
}

#Prepare the function extract_values to extract values of interest from the generated lists
extract_values_phot <- function(nested_list) {
  data.frame(intercept = nested_list$summary[[3]],
             slope = nested_list$summary[[4]],
             r2 = nested_list$summary[[5]],
             rate = nested_list$summary[[13]])
}

#Calculate photosynthesis rates
Pnet.df_rate <- Pnet.df %>%
  mutate(rate_lists = map(data, auto_nest_rate)) %>%
  mutate(rate_values = map(rate_lists, extract_values_phot)) %>%
  select(-rate_lists) %>%
  unnest(rate_values) %>%
  relocate(data, .after = rate)

#Find best fits for rates that are positive
Pnet.df_rate_pos <- Pnet.df_rate %>%
  group_by(Lineage) %>%
  filter(r2 == max(r2[rate > 0])) %>%
  mutate(rate = mean(rate)) %>% #If multiple rolling rates are selected due to the same r2 scores, keep the mean rate
  mutate(slope = mean(slope)) %>%
  mutate(intercept = mean(intercept)) %>%
  distinct()

#Find best fits for rates that are negative (for 21 and 31°C)
Pnet.df_rate_neg <- Pnet.df_rate %>%
  group_by(Lineage) %>%
  filter(r2 == max(r2[rate < 0])) %>%
  mutate(rate = mean(rate)) %>%
  mutate(slope = mean(slope)) %>%
  mutate(intercept = mean(intercept)) %>%
  distinct()

#Keep the best fits
Pnet.df_rate_max <- rbind(Pnet.df_rate_pos, Pnet.df_rate_neg) %>%
  group_by(Lineage) %>%
  filter(r2 == max(r2))
```

Calculate & adjust for background photosynthesis rates using blanks
```{r}
#Select blank measurements
Pnet.df_blank <- PR.df %>%
  filter(phase == 'light') %>%
  filter(Lineage == 'Blank') %>%
  filter(time > 1500) %>%
  filter(time < 2100) %>%
  filter(EE_treatment != 'TP1' | Replicate != 1) %>% #TP1 usually has 2 technical replicates, select either one
  select(!c(Replicate, Run, Chamber, phase)) %>%
  nest(data = c(time, oxygen))

#Calculate background photosynthesis rates
Pnet.df_blank_rate <- Pnet.df_blank %>%
  mutate(rate_lists = map(data, auto_nest_rate)) %>%
  mutate(rate_values = map(rate_lists, extract_values_phot)) %>%
  select(-rate_lists) %>%
  unnest(rate_values) %>%
  relocate(data, .after = rate)

Pnet.df_blank_rate_pos <- Pnet.df_blank_rate %>%
  group_by(EE_treatment) %>%
  filter(r2 == max(r2[rate > 0])) %>%
  mutate(rate = mean(rate)) %>%
  mutate(slope = mean(slope)) %>%
  mutate(intercept = mean(intercept)) %>%
  distinct()

Pnet.df_blank_rate_neg <- Pnet.df_blank_rate %>%
  group_by(EE_treatment) %>%
  filter(r2 == max(r2[rate < 0])) %>%
  mutate(rate = mean(rate)) %>%
  mutate(slope = mean(slope)) %>%
  mutate(intercept = mean(intercept)) %>%
  distinct()

Pnet.df_blank_rate_max <- rbind(Pnet.df_blank_rate_pos, Pnet.df_blank_rate_neg) %>%
  group_by(EE_treatment) %>%
  filter(r2 == max(r2))

#Adjust for background photosynthesis rates
Pnet.df_adj <- rbind(Pnet.df_rate_max, Pnet.df_blank_rate_max) %>%
  group_by(EE_treatment) %>%
  mutate(adj_rate = rate - rate[Lineage == 'Blank']) %>%
  relocate(data, .after = adj_rate)
```

Convert the adjusted photosynthesis rates to pmolO2/min/cell
```{r}
#Convert to pmolO2/min
Pnet.conversion <- convert_rate(Pnet.df_adj$adj_rate, 
                        oxy.unit = '%Air',           # oxygen unit to convert from
                        time.unit = 's',             # time unit to convert from
                        output.unit = 'pmol/min',    # unit to convert to
                        volume = 0.00146,            # chamber volume in litres (1.46 ml)
                        t = 21,                      # temperature, in °C (change for each temperature treatment)
                        S = 33,                      # salinity, in ppt
                        P = 1.013253)                # atmospheric pressure, in bar                   

Pnet.df_converted <- list(adj_rate = Pnet.conversion$summary[[13]], pmolO2.s = Pnet.conversion$summary[[29]]) %>%
  as.data.frame()

Pnet.df_final <- left_join(Pnet.df_adj, Pnet.df_converted, by = 'adj_rate') %>%
  distinct() %>%
  mutate(pmolO2.min = pmolO2.s*60) %>% #convert from pmol/s to pmol/min
  mutate(pmolO2.min.cell = pmolO2.min/(750000*1.46)) #normalise for cell density (750,000 cells/ml)
```

Visualize photosynthesis rates
```{r}
Pnet.plot <- Pnet.df_final %>%
  filter(Lineage != 'Blank') %>%
  ggplot(aes(y = pmolO2.min.cell, x = EE_treatment, fill = EE_treatment)) +
  geom_hline(aes(yintercept = 0), colour="black", size = 0.5, linetype = 'dashed') +
  geom_boxplot() +
  geom_point() +
  scale_x_discrete('Experimental evolution treatment') +
  scale_y_continuous(expression(P[net]~(pmol~O[2]~min^-1~cell^-1))) +
  scale_fill_manual(values=c("#8B0000","#E4B031", "#26727B", "#4770B3", "#58595b")) +
  theme_bw() +
  theme(panel.grid.major.x = element_blank(),
        panel.grid.minor = element_blank(),
        strip.text.x=element_text(size=14, face="bold"),
        strip.text.y=element_text(size=14, face="bold"),
        legend.title = element_text(size = 14, face = "bold"),
        legend.text = element_text(size = 12),
        axis.line = element_line(colour = "black"),
        axis.title.x = element_text(size = 14, colour = "black", face = "bold"),
        axis.text.x = element_text(size = 12, colour = "black"), 
        axis.title.y = element_text(size = 14, colour = "black", face = "bold"), 
        axis.text.y = element_text(size = 12, colour = "black")) +
        guides(fill = FALSE, colour = FALSE)
Pnet.plot
```

Save photosynthesis rates
```{r}
Pdf_xl <- Pdf_final %>%
  select(-data) %>%
  arrange(by_group = Lineage) %>%
  write_xlsx(path = "./Pnet rates - 21C.xlsx", #For each temperature, change file name
             col_names = TRUE,
             format_headers = TRUE)
```

Rd, Pnet and Pgross plots
```{r}
#Respiration plots
Rd.df_import <- read_xlsx("/Users/hugo/Documents/Melbourne PhD/Experimental work/Chapter 3/TPA/TPA - Rd & Pnet rates.xlsx", sheet = 1) %>%
  filter(Lineage != 'Blank') %>%
  mutate(Rate_type = 'respiration')

tukey_hsd.Rd.27C <- aov(pmolO2.min.cell ~ EE_treatment, data = Rd.df_import[Rd.df_import$Growth_temperature == 27, ]) %>%
  tukey_hsd() %>%
  filter(p.adj < 0.05)

tukey_hsd.Rd.29C <- aov(pmolO2.min.cell ~ EE_treatment, data = Rd.df_import[Rd.df_import$Growth_temperature == 29, ]) %>%
  tukey_hsd() %>%
  filter(p.adj < 0.05)

tukey_hsd.Rd.31C <- aov(pmolO2.min.cell ~ EE_treatment, data = Rd.df_import[Rd.df_import$Growth_temperature == 31, ]) %>%
  tukey_hsd() %>%
  filter(p.adj < 0.05)
#No significant differences in Rd at any temp (27-31°C)

Rd.boxplots <- Rd.df_import %>%
  filter(Growth_temperature >= 27) %>%
  ggplot(aes(y = pmolO2.min.cell, x = EE_treatment, fill = EE_treatment)) +
  geom_hline(aes(yintercept = 0), colour="black", size = 0.5, linetype = 'dashed') +
  geom_boxplot() +
  facet_grid(.~Growth_temperature, labeller = labeller(Growth_temperature=temperature.labs)) +
  scale_x_discrete(labels = treatment.labs, 'Experimental evolution treatment') +
  scale_y_continuous(expression(bold(R[d]~(pmol~O[2]~min^-1~cell^-1)))) +
  scale_fill_manual(values=c("#8B0000","#E4B031", "#26727B", "#4770B3", "#58595b")) +
  theme_bw() +
  theme(panel.grid.major.x = element_blank(),
        panel.grid.minor = element_blank(),
        strip.text.x=element_text(size=14, face="bold"),
        strip.text.y=element_text(size=14, face="bold"),
        legend.title = element_text(size = 14, face = "bold"),
        legend.text = element_text(size = 12),
        axis.line = element_line(colour = "black"),
        axis.title.x = element_text(size = 14, colour = "black", face = "bold"),
        axis.text.x = element_text(size = 12, angle = 45, hjust = 1, colour = "black"), 
        axis.title.y = element_text(size = 14, colour = "black", face = "bold"), 
        axis.text.y = element_text(size = 12, colour = "black")) +
        guides(fill = FALSE, colour = FALSE)
Rd.boxplots

#Net photosynthesis plots
Pnet.df_import <- read_xlsx("/Users/hugo/Documents/Melbourne PhD/Experimental work/Chapter 3/TPA/TPA - Rd & Pnet rates.xlsx", sheet = 2) %>%
  filter(Lineage != 'Blank') %>%
  mutate(Rate_type = 'photosynthesis') %>%
  filter(!(Growth_temperature == 31 & Lineage %in% c('TP1-3', 'TP2-2', 'TP3-2', 'TP4-3', 'TP5-1'))) %>%
  filter(!(Growth_temperature == 29 & Lineage %in% c('TP1-2', 'TP2-1', 'TP4-3', 'TP4-5', 'TP5-1', 'TP5-5'))) %>%
  filter(!(Growth_temperature == 27 & Lineage %in% c('TP2-6')))

tukey_hsd.Pnet.27C <- aov(pmolO2.min.cell ~ EE_treatment, data = Pnet.df_import[Pnet.df_import$Growth_temperature == 27, ]) %>%
  tukey_hsd() %>%
  filter(p.adj < 0.05) %>%
  mutate(Growth_temperature = 27)

tukey_hsd.Pnet.29C <- aov(pmolO2.min.cell ~ EE_treatment, data = Pnet.df_import[Pnet.df_import$Growth_temperature == 29, ]) %>%
  tukey_hsd() %>%
  filter(p.adj < 0.05) %>%
  mutate(Growth_temperature = 29)

tukey_hsd.Pnet.31C <- aov(pmolO2.min.cell ~ EE_treatment, data = Pnet.df_import[Pnet.df_import$Growth_temperature == 31, ]) %>%
  tukey_hsd() %>%
  filter(p.adj < 0.05) %>%
  mutate(Growth_temperature = 31)

tukey_hsd.Pnet <- rbind(tukey_hsd.Pnet.27C, tukey_hsd.Pnet.29C) %>%
  rbind(tukey_hsd.Pnet.31C)

Pnet.boxplots <- Pnet.df_import %>%
  filter(Growth_temperature >= 27) %>%
  ggplot() +
  geom_hline(aes(yintercept = 0), colour="black", size = 0.5, linetype = 'dashed') +
  geom_boxplot(aes(y = pmolO2.min.cell, x = EE_treatment, fill = EE_treatment)) +
  stat_pvalue_manual(tukey_hsd.Pnet, label = "p.adj.signif", y.position = c(0.36, 0.33), tip.length = 0.01) +
  facet_grid(.~Growth_temperature, labeller = labeller(Growth_temperature=temperature.labs)) +
  scale_x_discrete(labels = treatment.labs, 'Experimental evolution treatment') +
  scale_y_continuous(expression(bold(P[net]~(pmol~O[2]~min^-1~cell^-1)))) +
  scale_fill_manual(values=c("#8B0000","#E4B031", "#26727B", "#4770B3", "#58595b")) +
  theme_bw() +
  theme(panel.grid.major.x = element_blank(),
        panel.grid.minor = element_blank(),
        strip.text.x=element_text(size=14, face="bold"),
        strip.text.y=element_text(size=14, face="bold"),
        legend.title = element_text(size = 14, face = "bold"),
        legend.text = element_text(size = 12),
        axis.line = element_line(colour = "black"),
        axis.title.x = element_text(size = 14, colour = "black", face = "bold"),
        axis.text.x = element_text(size = 12, angle = 45, hjust = 1, colour = "black"), 
        axis.title.y = element_text(size = 14, colour = "black", face = "bold"), 
        axis.text.y = element_text(size = 12, colour = "black")) +
        guides(fill = FALSE, colour = FALSE)
Pnet.boxplots

#Gross photosynthesis plots
library(plyr)
Rd.df_import2 <- match_df(Rd.df_import, Pnet.df_import, on = c('Growth_temperature', 'EE_treatment', 'Lineage'))
detach("package:plyr", unload=TRUE)

Pgross.df <- rbind(Rd.df_import2, Pnet.df_import) %>%
  group_by(Growth_temperature, Lineage) %>%
  mutate(Pgross = pmolO2.min.cell[Rate_type == 'photosynthesis'] + pmolO2.min.cell[Rate_type == 'respiration']) %>%
  ungroup() %>%
  dplyr::select(Growth_temperature, EE_treatment, Lineage, Pgross) %>%
  distinct()

tukey_hsd.Pgross.27C <- aov(Pgross ~ EE_treatment, data = Pgross.df[Pgross.df$Growth_temperature == 27, ]) %>%
  tukey_hsd() %>%
  filter(p.adj < 0.05) %>%
  mutate(Growth_temperature = 27)

tukey_hsd.Pgross.29C <- aov(Pgross ~ EE_treatment, data = Pgross.df[Pgross.df$Growth_temperature == 29, ]) %>%
  tukey_hsd() %>%
  filter(p.adj < 0.05) %>%
  mutate(Growth_temperature = 29)

tukey_hsd.Pgross.31C <- aov(Pgross ~ EE_treatment, data = Pgross.df[Pgross.df$Growth_temperature == 31, ]) %>%
  tukey_hsd() %>%
  filter(p.adj < 0.05) %>%
  mutate(Growth_temperature = 31)

tukey_hsd.Pgross <- rbind(tukey_hsd.Pgross.27C, tukey_hsd.Pgross.29C) %>%
  rbind(tukey_hsd.Pgross.31C)

Pgross.boxplots <- Pgross.df %>%
  filter(Growth_temperature >= 27) %>%
  ggplot() +
  geom_hline(aes(yintercept = 0), colour="black", size = 0.5, linetype = 'dashed') +
  geom_boxplot(aes(y = Pgross, x = EE_treatment, fill = EE_treatment)) +
  stat_pvalue_manual(tukey_hsd.Pgross, label = "p.adj.signif", y.position = c(0.26, 0.22), tip.length = 0.01) +
  facet_grid(.~Growth_temperature, labeller = labeller(Growth_temperature=temperature.labs)) +
  scale_x_discrete(labels = treatment.labs, 'Experimental evolution treatment') +
  scale_y_continuous(expression(bold(P[gross]~(pmol~O[2]~min^-1~cell^-1)))) +
  scale_fill_manual(values=c("#8B0000","#E4B031", "#26727B", "#4770B3", "#58595b")) +
  theme_bw() +
  theme(panel.grid.major.x = element_blank(),
        panel.grid.minor = element_blank(),
        strip.text.x=element_text(size=14, face="bold"),
        strip.text.y=element_text(size=14, face="bold"),
        legend.title = element_text(size = 14, face = "bold"),
        legend.text = element_text(size = 12),
        axis.line = element_line(colour = "black"),
        axis.title.x = element_text(size = 14, colour = "black", face = "bold"),
        axis.text.x = element_text(size = 12, angle = 45, hjust = 1, colour = "black"), 
        axis.title.y = element_text(size = 14, colour = "black", face = "bold"), 
        axis.text.y = element_text(size = 12, colour = "black")) +
        guides(fill = FALSE, colour = FALSE)
Pgross.boxplots

#Plot Rd and Pnet together
Rd.df_import3 <- Rd.df_import %>%
  group_by(Growth_temperature, EE_treatment) %>%
  mutate(mean_Rd = mean(pmolO2.min.cell)) %>%
  mutate(sd_Rd = sd(pmolO2.min.cell)) %>%
  ungroup() %>%
  dplyr::select(Growth_temperature, EE_treatment, mean_Rd, sd_Rd) %>%
  distinct()

Pnet.df_import2 <- Pnet.df_import %>%
  group_by(Growth_temperature, EE_treatment) %>%
  mutate(mean_Pnet = mean(pmolO2.min.cell)) %>%
  mutate(sd_Pnet = sd(pmolO2.min.cell)) %>%
  ungroup() %>%
  dplyr::select(Growth_temperature, EE_treatment, mean_Pnet, sd_Pnet) %>%
  distinct()

RdPnet.df <- left_join(Rd.df_import3, Pnet.df_import2, by = c('Growth_temperature', 'EE_treatment')) 

RdPnet.plot <- RdPnet.df %>%
  filter(Growth_temperature >= 27) %>%
  mutate(Growth_temperature = as.factor(Growth_temperature)) %>%
  ggplot(aes(x = abs(mean_Rd), y = mean_Pnet, fill = EE_treatment, group = interaction(Growth_temperature, EE_treatment))) +
  geom_linerange(aes(xmin = abs(mean_Rd) - sd_Rd, xmax = abs(mean_Rd) + sd_Rd, colour = EE_treatment), size = 0.4) +
  geom_linerange(aes(ymin = mean_Pnet - sd_Pnet, ymax = mean_Pnet + sd_Pnet, colour = EE_treatment), size = 0.4) +
  geom_abline(intercept = 0, slope = 1, colour="black", size = 0.5, linetype = 'dashed') +
  geom_point(pch = 21, size = 2) +
  facet_grid(.~Growth_temperature, labeller = labeller(Growth_temperature=temperature.labs)) +
  scale_x_continuous(expression(bold(R[d]~(pmol~O[2]~min^-1~cell^-1))), limits = c(-0.1, 0.5), expand = c(0,0)) +
  scale_y_continuous(expression(bold(P[net]~(pmol~O[2]~min^-1~cell^-1))), limits = c(-0.1, 0.5), expand = c(0,0)) +
  scale_colour_manual(labels = treatment.labs, values=c("#8B0000","#E4B031", "#26727B", "#4770B3", "#58595b")) +
  scale_fill_manual('Experimental evolution treatment', labels = treatment.labs,
                    values=c("#8B0000","#E4B031", "#26727B", "#4770B3", "#58595b")) +
  theme_bw() +
  theme(panel.grid.major.x = element_blank(),
        panel.grid.minor = element_blank(),
        strip.text.x=element_text(size=14, face="bold"),
        strip.text.y=element_text(size=14, face="bold"),
        legend.title = element_text(size = 14, face = "bold"),
        legend.text = element_text(size = 12),
        legend.position = 'top', 
        axis.line = element_line(colour = "black"),
        axis.title.x = element_text(size = 14, colour = "black", face = "bold"),
        axis.text.x = element_text(size = 12, angle = 45, hjust = 1, colour = "black"), 
        axis.title.y = element_text(size = 14, colour = "black", face = "bold"), 
        axis.text.y = element_text(size = 12, colour = "black")) +
  guides(colour = F)
RdPnet.plot
```

Export plots with Fv/Fm output
```{r}
PR.boxplots <- ggarrange(Rd.boxplots + rremove('xlab'),
                         Pnet.boxplots + rremove('xlab'),
                         Pgross.boxplots + rremove('xlab'),
                         common.legend = TRUE,
                         labels = c("A", "B", "C"),
                         heights = c(1, 1, 1),
                         widths = c(1, 1, 1),
                         ncol = 1,
                         align = "v") 
PR.boxplots

PR.boxplots.annot <- annotate_figure(PR.boxplots,
                                     bottom = text_grob("Experimental evolution treatment", color = "black", face = "bold", size = 14))
PR.boxplots.annot

PR.FvFm.plots <- ggarrange(RdPnet.plot,
                           FvFm.plots,
                           heights = c(1, 1),
                           labels = "A",
                           ncol = 1,
                           align = "hv") 
PR.FvFm.plots

PR.FvFm.plots.annot <- annotate_figure(PR.FvFm.plots,
                                       bottom = text_grob("Experimental evolution treatment", color = "black", face = "bold", size = 14))
PR.FvFm.plots.annot
```

Save plots
```{r}
ggsave(plot = PR.boxplots.annot, "TPA - PR boxplots.pdf", 
       path = "./",
       width = 7.5,
       height = 12.5,
       units = 'in',
       dpi = 300,
       device = cairo_pdf)

ggsave(plot = PR.FvFm.plots.annot, "TPA - PR with FvFm.pdf", 
       path = "./",
       width = 10,
       height = 10,
       units = 'in',
       dpi = 300,
       device = cairo_pdf)
```
